---
title: "hw4markdown"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
I am going to load in my libraries and then import data on the real GDP growth data from 1947 to present day(the longest sample available) via FRED.
```{r}
library(readxl)
library(ggplot2)
GDPC1_percentchange <- read_excel("C:/Econometrics/ECONOMETRICS 674/GDPC1-percentchange.xls", 
    skip = 10)
GDPC1_percentchange<- data.frame(GDPC1_percentchange)

```
After importing the data, I will plot with a line chart. 
```{r}
plot(GDPC1_percentchange, type = "l",
     main="Real GDP Growth Data", xlab= "Date of Observation", ylab = "Percent Change of GDP per Quarter")

```
Another way to get this data and plot it is through the quantmod and Quandl packages, which will upload the data automatically. This is demonstrated below. 
```{r}
library(quantmod)
library(Quandl)
GDP_percentchange_quandl <- Quandl("FRED/GDPC1", start_date="1947-04-01", end_date="2021-10-01", transform = "rdiff", collapse = "quarterly")
plot(GDP_percentchange_quandl, type = "l",
     main="Real GDP Growth Data", xlab= "Date of Observation", ylab = "Percent Change of GDP per Quarter")

```

Now that I have constructed the same plot using three different methods(Excel, importing file into R, and Quandl), I can compare the three plots. These three plots look identical. Now I will find the mean and variance of the data. 
```{r}
mean1<-mean(GDPC1_percentchange$GDPC1_PCH)
var1<-var(GDPC1_percentchange$GDPC1_PCH)
```
From this, it can be seen that the mean of the data is .7707 and the variance is 1.344. Now I will predict real GDP growth with its unconditional mean, which would be .7707. If I were to treat the GDP growth distribution as Normal, what would be the probability of real GDP growth being negative?
```{r}
sd1<- sqrt(var1)
pnorm1<- pnorm(0, mean1, sd1)
print(pnorm1)
```

This means that the probability of real GDP growth being negative is .2530 or 25.30%. In addition, the probability is 19% that real GDP growth will be below one standard deviation from the mean, which is shown by the code below.
```{r}
pnorm(mean1-sqrt(var1), mean1,var1)
```

I will now use the unconditional distribution under the assumption of Normality to construct the 70% and 95% prediction intervals around the mean.
```{r}
predint1<- qnorm(.95, mean = mean1, sd=sd1)
predint2<- qnorm(.70, mean = mean1, sd=sd1)
```
The 95% prediction interval around the mean is 2.678 and the 70% prediction interval around the mean is 1.379. This means that 95% of the values will lie below 2.678 and 70% of the values will lie below 1.379.
